### 2017-06-27 paper/coverletter.tex
Add coverletter

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.
%Version for spanish users, by dgarhdez

\documentclass[12pt,letter,roman]{moderncv}

% moderncv themes
% style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvstyle{classic}
% color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\moderncvcolor{green}
% to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\renewcommand{\familydefault}{\sfdefault}
% uncomment to suppress automatic page numbering for CVs longer than one page
%\nopagenumbers{}

% character encoding
\usepackage[utf8]{inputenc}

% adjust the page margins
\usepackage[scale=0.75]{geometry}

% personal data
\name{Semyeong}{Oh}
\title{Resumé title}
\address{4 Ivy Ln}{Princeton, NJ 08540}{United States}
\email{semyeong@astro.princeton.edu}

\begin{document}

% recipient data
\recipient{Dr. Leslie Sage}{968 National Press Building\\529 14th Street NW\\Washington DC 20045-1938\\United States}
\date{\today}
\opening{Dear Dr. Sage,}
\closing{Best regards,}
\makelettertitle

We submit the enclosed document for consideration for a Nature Letter.
The paper lays out a simple and clear account for evidences in stellar
chemical abundances that massive rocky accretion happened in one of the two
stars that are comoving in space, and possibly also in the other one albeit in
smaller amount.
We think the result is one of the most convincing examples that shows a
catastrophic ending possible to planets around sun-like stars, and thus will be
interesting to scientists in other disciplines as well as those in astronomy or
exoplanet field.

The document has been formatted with \LaTeX\ article class.
The summary is 286 words long, and the main text has $\approx 1424$ words,
29 references,
and two single-panel figures and one table, which should fit the page limit
of a Nature Letter.
We also include six extended data figure that will help the interested readers
get the full context, as well as supplementary material discussing several
aspects of the main text.

\vspace{0.5cm}

\makeletterclosing

\end{document}